---
output:
  word_document: default
  html_document: default
---
------------------------------------------------------------------------

title: "LAB 2" author: "Nehal Ur Rahman" date: "`r Sys.Date()`" output: word_document

#NAME: NEHAL UR RAHMAN
#STUDENT ID: 991691259

##Introduction 
#In Lab 2 We will be analyzing a College dataset and performing various functions to understand the variables. We will also be creating a new variable and provide visualizations.


```{r}
#Import
#Here we are going to load and read all variables with its values from the college1 dataset.
college <- read.csv("college1.csv")
head(college)

```

```{r}
#The fix function is used to fix the 1st column and not store it as data as they are just labels
fix(college)
```

```{r}
#Here we fix the dataset by adding a column called row.names which records the name of all the universities
row.names(college) = college[,1]
fix(college)
```

```{r}
#Using the function given below we delete the 1st column in the college dataset as it is not required in our analysis.
college = college[,-1]
fix(college)

```


```{r}
#The as.factor function converts the character variable(Private) to vector with numerical values
college$Private<-as.factor(college$Private)
```

###Question 1
```{r}

#First we create a variable by using the function rep() which replicates the college with 777 number of rows with a value of "No" 
Elite <- rep("No",nrow(college))
#Now we record the values as "Yes" in the Elite column with a condition that the proportion of students coming from the top 10% of their high school exceeds 50%. 
Elite[college$Top10perc >50] <- "Yes"
#The as.factor function displays the variable(Elite) as vector with levels(Yes & No)
Elite <- as.factor(Elite)
#Now we create a dataframe with college and elite
college <- data.frame(college , Elite)
TotalEliteSchools <- length(college$Elite[college$Elite=="Yes"])
#We then calculate the total number of Elite colleges and display the number. 
message("The total number of Elite Schools are : ", TotalEliteSchools)

```
###Question 2
```{r}
#The summary function is used here to get details of the Elite column
summary(college$Elite)
```
```{r}
#A side by side boxplot of Outstate Vs Elite is created using plot function
plot(college$Outstate ~ college$Elite, col = c("violet", "orange"), xlab="ELITE",ylab="OUTSTATE", main = "Outstate vs Elite", border = "black")
```
#Boxplot: From the boxplot we can see that the number of Elite colleges are more in the Outstate.

###Question 3
```{r}
#The hist() function is used here to produce histograms with variable number of bins for 3 of the quantitative variables like Top10perc, Top25perc and Grad.Rate.
#breaks = 6 & 8 is assigned first which gives lesser number of bins
#breaks = 12 is assigned next to get more number of bins. 

#The par() function divides the frame into the required number to display the histograms within one window.
par(mfcol=c(2,3))

hist(college$Top10perc, col = "blue",breaks=8, xlab = "Top 10%", ylab = "Value", main="Students from Top 10% of H.S")
hist(college$Top10perc, col = "green",breaks=12, xlab = "Top 10%", ylab = "Value", main="Students from Top 10% of H.S")

hist(college$Top25perc, col = "blue",breaks= 6, xlab = "Top 25%", ylab = "Value", main="Students from Top 25% of H.S")
hist(college$Top25perc, col = "green",breaks=12, xlab = "Top 25%", ylab = "Value", main="Students from Top 25% of H.S")

hist(college$Grad.Rate, col = "blue",breaks=6, xlab = "Graduation rate", ylab = "Value", main="College Graduation Rate")
hist(college$Grad.Rate, col = "green",breaks=12, xlab = "Graduation rate", ylab = "Value", main="College Graduation Rate")

```
#The histogram for the 3 variables: Top10perc, Top25perc and Grad.Rate have been displayed with different number of bins.
